---
title: "ESM206_lab5"
author: "Kelsie Fronheiser"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 5 Objectives

- Learn to parse dates with 'lubridate'
- Get counts of observations with count()
- Uncount() rows
- Crate heat map with geom_tile()

```{r, include =FALSE}
# Read in data and attach packages

library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data","lobster_abundance.csv"),
                              na = "-99999") %>% 
  janitor::clean_names()
```

Use tidyr::uncount() function to convert lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% tidyr::uncount(lobster_count)
```

### Exploratory data visualization

We're only going to consider "site" as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales =  'free')

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)

# by looking at the histograms and quantile quantile plots we can determine the data is noramlly distributed

```

### Convert the 'date' column to class 'Date'

Use the 'lubridate' package to convert to Date format, and then to help us easily parse month and year

```{r}

lobster_date <-  lobster_tidy %>% 
  mutate(
    date_new = lubridate:: mdy(date)
  )
```

Now let's parse year and month using libridate::month() and lubridate::year():

```{r}
lobster_parse_date <-  lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE), 
    obs_year = lubridate::year(date_new)
  )
```

Now let's find counts of observed lobsters based on different hierachial groupings using count():


First we will count lobsters by year and month, then by year, then by site

```{r}
lobster_ym <-  lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_ym

lobster_y <-  lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_y

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)

lobster_site
```

If we want to creat a summary table that contains stats other than count by group, it is eassier to use group_by + n()
```{r}
lobster_summary <-  lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )
    
    lobster_summary
```

### Find confidence intervals

Use t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples (will be coved in lec week 6)

```{r}
ivee_lobsters <-  lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)
```
### Two- sample t-test to compare means

Is there a sig diff between lobster sizes between two sites (Naples & Mohawk)?

We've done our neccessary exploratory analyses to dtermine that a 2-sample t-test for means comparfison is appropriate

Null hypo.is default that says all pops/samples have same mean, not sig diff

```{r}
napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)

mn_ttest
```

There is a sig diff in lobster size between Naples and Mohawk Reef (t(df) = statistic, p < 0.001, alpha = 0.5). 

p value is not the most important thing to communicate to audience

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <-  t.test(size_mm ~ site, data= lobster_mn)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic
```

Drawn from pops. w/ difference in means
There is a sig diff (t(1850.8) = -19.849, p = 2.2e-16).

Here is the p-value: 'r mn_ttest2$p.value'

Here is my t-statistic: 'r mn_ttest2$statistic'

### Now a heatmap
```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```



























